### 2024-06-12 paper/paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{agujournal2019}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={From Precipitation to Prediction: Using ConvLSTM Models for Comprehensive River Runoff Forecasting},
  pdfauthor={Florian Börgel; Sven Karsten; Karoline Rummel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\journalname{Climate Dynamics}

\draftfalse

\begin{document}
\title{From Precipitation to Prediction: Using ConvLSTM Models for
Comprehensive River Runoff Forecasting}

\authors{Florian Börgel\affil{1}, Sven Karsten\affil{1}, Karoline
Rummel\affil{1}}
\affiliation{1}{Leibniz-Institute for Baltic Sea Research Warnemünde, }
\correspondingauthor{Florian Börgel}{florian.boergel@io-warnemuende.de}

\begin{abstract}
a
\end{abstract}

\section*{Plain Language Summary}
b

\section{Introduction}\label{introduction}

River runoff is an important component of the global water cycle as it
comprises about one third of the precipitation over land areas (Hagemann
et al., 2020). Therefore, accurate runoff forecasting is crucial for
effective water resources management, particularly over extended periods
(Fang et al., 2019; Tan et al., 2018). In the context of climate change
studies, river runoff is usually generated in two ways. First, river
runoff as input for ocean models can be created using hydrological
models such as the Hydrological Discharge (HD) model (Hagemann et al.,
2020). HD models calculate the water balance using hydrological
processes (e.g.~snow, glaciers, soil moisture, groundwater
contribution). It represents a complex forecasting tool that uses
underlying physical processes. The second approach uses data-driven
models that integrate statistical correction, using land surface schemes
of global or regional climate models {[}\texttt{missing}{]}.

With the recent rise of machine learning in climate research various
model architectures have also been tested for river runoff forecasting.
Common approaches employ feed-forward artificial neural networks,
support vector machines, adaptive neuro-fuzzy inference systems, and
notably, Long Short-Term Memory (LSTM) neural networks that have gained
traction for long-term hydrological forecasting due to their excellent
performance (\textbf{humphrey2016hybrid?}; \textbf{huang2014monthly?};
\textbf{ashrafi2017fully?}; \textbf{liu2020streamflow?};
\textbf{fang2022application?}; \textbf{kratzert2018rainfall?}).

LSTM networks, first introduced by (\textbf{hochreiter1997long?}) , are
an evolution of the classical Recurrent Neural Networks (RNNs). A
significant advantage of this architecture is the memory cell's ability
to retain gradients. This mechanism addresses the vanishing gradient
problem, where, as input sequences elongate, the influence of initial
stages becomes harder to capture, causing gradients of early input
points to approach zero. LSTMs have shown stability and efficacy in
sequence-to-sequence predictions. This performance in modeling
long-range dependencies has been validated in various studies
{[}\texttt{missing}{]}.

However, a limitation of LSTMs is their inability to effectively capture
two-dimensional structures, an area where Convolutional Neural Networks
(CNNs) excel. Recognizing this limitation, SHI et al. (2015) proposed a
Convolutional LSTM (ConvLSTM) architecture, which combines the strengths
of both LSTM and CNN. The ConvLSTM network has been proven useful for
spatio-temporal applications such as precipitation nowcasting (SHI et
al., 2015), flood forecasting (\textbf{moishin2021designing?}), and
river runoff forecasting (\textbf{ha2021?};
\textbf{zhu2023spatiotemporal?}).

In this work, we demonstrate that ConvLSTM networks are a reliable
method for predicting multiple rivers simultaneously, using only
atmospheric forcing, even in the absence of a fully functional
hydrological model with a complex parameterization. Still, data of a
suitable HD model is used as reference to train the network. We use the
Baltic Sea catchment as an example to illustrate our approach. Although
the methodology we propose is universally applicable across various
geographic regions, the Baltic Sea represents a challenging region due
to its unique hydrological characteristics, being nearly decoupled from
the open ocean (see Figure). As a consequence, the salinity of the
Baltic Sea is driven to a large part by freshwater supply from rivers.
Freshwater enters the Baltic Sea through river runoff or positive net
precipitation (precipitation minus evaporation) over the sea surface.
The net precipitation accounts for 11 \(\%\) and the river input for 89
\(\%\) of the total freshwater input (\textbf{meier2002simulated?}).
Modelling the Baltic Sea is therefore to a large part the result of the
quality of the river input, that is used for the simulation. This makes
the accurate modeling of river runoff especially critical for
simulations pertaining to the Baltic Sea.

The main focus of this work is to present a ConvLSTM architecture that
is able to predict daily river runoff for 97 rivers across the Baltic
Sea catchment (Section~\ref{sec-methodology}). To train the network, we
use data from the E-HYPE HD model {[}\texttt{missing}{]} as reference
output data and data from the UERRA project (Uncertainties in Ensembles
of Regional Re-Analyses,
\href{http://www.uerra.eu/}{http://www.uerra.eu/)},) as atmospheric
forcing (Section~\ref{sec-technical_details}). The quality of the model
is evaluated in \textbf{?@sec-sec-evaluation} and subsequently the
trained ConvLSTM is used in combination with the MOM5 ocean model for
the Baltic Sea (Neumann et al., 2021)
(Section~\ref{sec-MOM-application}). The obtained results are further
discussed in Section~\ref{sec-discussion} and concluded in
Section~\ref{sec-conclusions}.

\section{Implemented model architecture}\label{sec-methodology}

\subsection{The main idea}\label{sec-main_idea}

We assume that the runoff at a specific point in time \(t\) for all
\(N_r\) considered rivers, collected in the vector
\(\vec{R}^t \in \mathbb{R}^{N_r}\), can be accurately approximated by a
functional \(\vec{M}(\{X^t_k[x,y,\tau]\})\) of \(k=1,\ldots N_k\)
atmospheric fields \(X^t_k[x,y,\tau]\) which are known for the past
\(\tau=1,\ldots N_\tau\) time instances. This relationship is expressed
as:

\[
\begin{aligned}
\vec{R}^t = \vec{M}(\{X^t_k[x,y,\tau]\}) \ .
\end{aligned}
\]

The atmospheric fields are evaluated over a spatial domain
\(x=1,\ldots N_x\) and \(y=1,\ldots N_y\) which is sufficiently large to
capture all significant local and non-local contributions of the
atmospheric fields to the river runoff.

Typically, such a mapping is realized using a hydrological model that
simulates all relevant physical processes, transforming variables like
precipitation and evaporation into river runoff. This process relies
heavily on domain knowledge to tune all parameters to reasonable values.

As an alternative, we propose that this functional can be adequately
represented by a combination of a convolutional Long-Short Term Memory
(ConvLSTM) model with a subsequent fully connected (FC) neural network.
This approach eliminates the need for detailed knowledge of the involved
processes and their modeling. Instead, these features can be ``learned''
by the network in an automated manner, i.e.~all free parameters are
optimized such that the network's output reproduces best the HD model
data with given atmospheric input fields. Our proposed network
architecture is visualized in Figure~\ref{fig-ConvLSTM_FC} and described
in detail in the following sections. To provide an overview, we will
discuss the main components of this architecture one-by-one.

\begin{figure}

\centering{

\includegraphics{./ConvLSTM_FC.png}

}

\caption{\label{fig-ConvLSTM_FC}\textbf{Combined ConvLSTM and FC network
architecture}. The starting point is the continuous time series of input
data \(\{\vec{X}^t\}\), upper white block. From this series, a
contiguous sequence of \(N_\tau\) elements (yellow block) is used to
feed a chain of \(N_\tau\) connected ConvLSTM cells (light blue block)
building the ConvLSTM network (grey block). The input sequence is mapped
via weighting matrices \(\pmb{\mathcal{M}}^g\) (green blocks) onto gate
vectors \(\vec{g}^t[\tau]\). The gate vectors are then used to update
the cell state \(\vec{C}^t[\tau-1]\) and the hidden state
\(\vec{H}^t[\tau-1]\) of the last ConvLSTM cell to the current values
\(\vec{C}^t[\tau]\) and \(\vec{H}^t[\tau]\), respectively. The update is
performed with the LSTM core equations collectively described by the
mapping \(L\), see Equation~\ref{eq-LSTM}. The weighting matrices
\(\pmb{\mathcal{N}}^g\) (green blocks) control how much of the last
hidden state enters the updated state. The final output of the ConvLSTM
\(\vec{H}^t[\tau]\) is then propagated to a FC network, which itself is
a chain of three FC layers consisting of weighting matrices
\(\pmb{\mathcal{W}}\) and connected via ReLU functions, see
Section~\ref{sec-FC_layer}. The final result is the river runoff
\(\vec{R}^t\) for all rivers considered at the current time instance
\(t\) (white block on the lower left). Note that all bias vectors are
omitted for the sake of clarity. See text for more information.}

\end{figure}%

\subsection{The ConvLSTM network}\label{the-convlstm-network}

\subsubsection{The LSTM approach}\label{sec-LSTM}

Before turning directly to the ConvLSTM the simpler architecture of the
plain Long-Short Term Memory (LSTM) model is examined which serves as a
foundation for understanding the more complex ConvLSTM.

The LSTM, a specialized form of Recurrent Neural Networks (RNNs), is
specifically designed to model temporal sequences
\(\vec{X}^t[1], \ldots \vec{X}^t[\tau],\ldots \vec{X}^t[N_\tau]\) of
\(N_\tau\) input quantities
\(\vec{X}^t[\tau] = (X^t_k[\tau]) \in \mathbb{R}^{N_k}\). This sequence
is taken from a dataset given in form of a time series \(\{\vec{X}^t\}\)
with the endpoint coinciding with the specific element in the time
series connected to time \(t\),
i.e.~\(\vec{X}^t[N_\tau] \equiv \vec{X}^t\), see
Figure~\ref{fig-ConvLSTM_FC}. Here, \(N_k\) represents the number of
input ``channels,'' which can correspond to different measurable
quantities. The LSTM's unique design allows it to adeptly handle
long-range dependencies, setting it apart from traditional RNNs in terms
of accuracy (see Figure~\ref{fig-lstm}).

\begin{figure}

\centering{

\includegraphics[width=10.5cm,height=\textheight]{./LSTM.png}

}

\caption{\label{fig-lstm}\textbf{Inner structure of a Long Short-Term
Memory Cell}. See Figure~\ref{fig-ConvLSTM_FC} and text for
information.}

\end{figure}%

The key component of the LSTM's innovation is its cell state,
\(\vec{C}^t[\tau] = (C^t_h[\tau]) \in \mathbb{R}^{N_h}\), which stores
state information, also referred to as \emph{long-term memory}. This
state information complements the so-called hidden state vector
\(\vec{H}^t[\tau] = (H^t_h[\tau]) \in \mathbb{R}^{N_h}\), which is also
known from simpler neural network architectures. In case of the LSTM,
the hidden state vector plays the role of the \emph{short-term memory}.
The cell state and the hidden state are vectors, where each element is
associated with one of the \(N_h\) hidden layers, labeled by \(h\),
which are internal, \emph{artificial} degrees of freedom that enable the
high adaptability of neural networks. These two state vectors are
determined through several self-parameterized gates, all in the same
vector space as \(\vec{C}^t[\tau]\), see Figure~\ref{fig-lstm} for a
visualization.

In particular, the forget gate \(\vec{F}^t[\tau]\) defines the portion
of the previous (long-term memory) cell state \(\vec{C}^t[\tau-1]\) that
should be kept, see dashed cyan box therein\hspace{0pt}. The input gate
\(\vec{I}^t[\tau]\) controls the contribution of the current input used
to update the long-term memory, \(\vec{C}^t[\tau]\) (magenta and yellow
boxes). The output gate, \(\vec{O}^t[\tau]\), then determines how much
of this updated long-term memory contributes to the new (short-term
memory) hidden state, \(\vec{H}^t[\tau]\) (black dashed box).

For a fixed point \(\tau\) in the sequence, the action of a LSTM cell,
i.e.~the connection between the input \(\vec{X}^t[\tau]\), the various
gates and the state vectors, is mathematically given as follows.

First, the elements of the input sequence together with the hidden state
are mapped onto auxiliary gate vectors, collectively denoted by
\(\vec{g}^t[\tau] = (g^t_h[\tau]) \in \mathbb{R}^{N_h}\), via \[
\begin{aligned}
g_h^t[\tau] =  \mathcal{M}^{g}_{hk} \, X^t_k[\tau] +  \mathcal{N}^{g}_{hh'} \, H^t_{h'}[\tau-1] + \mathcal{B}^g_{h} \ ,
\end{aligned}
\]

where \(g=i,f,o,c\) stands for the input, forget, output and cell-state
gate, respectively and Einstein's summation convention is employed,
i.e.~indices that appear twice are summed over. The calligraphic symbols
\(\mathcal{M}^{g}_{hk}, \mathcal{N}^{g}_{hh'}\) and
\(\mathcal{B}^g_{h}\) are the free parameters of the network that are
optimized for the given problem during the training, which is at the
heart of the any machine learning approach. The matrix
\(\pmb{\mathcal{M}}^{g} = (\mathcal{M}^{g}_{hk}) \in \mathbb{R}^{N_h \times N_k}\)
can be interpreted as a Markovian-like contribution of the current input
\(\vec{X}^t[\tau]\) to the gates, whereas the
\(\pmb{\mathcal{N}}^{g} = (\mathcal{N}^{g}_{hh'}) \in \mathbb{R}^{N_h \times N_h}\)
scales a non-Markovian part determined by the hidden state of the last
sequence point \(\tau-1\). The vector
\(\vec{\mathcal{B}}^g = (\mathcal{B}^g_{h}) \in \mathbb{R}^{N_h}\) is a
learnable bias. It should be stressed that these parameters do neither
depend on \(t\) nor on \(\tau\) and are thus optimized once for the
complete dataset \(\{\vec{X}^t\}\).

Note that this mapping is sometimes extended by a contribution to the
\(g_h^t[\tau]\) from the past cell state \(\vec{C}^t[\tau-1]\)
{[}\texttt{missing}{]}. Nevertheless, this mechanism called ``peeping''
is not further considered in this work.

For the sake of brevity, we can write the mapping more compactly in
matrix-vector form as
\begin{equation}\phantomsection\label{eq-LSTM-gates}{
\begin{aligned}
\vec{g}^t[\tau] = \pmb{\mathcal{M}}^{g}\vec{X}^t[\tau] + \pmb{\mathcal{N}}^{g}  \vec{H}^t[ \tau-1]+ \vec{\mathcal{B}}^g \,
\end{aligned}
}\end{equation}

Second, the actual gate vectors are computed by the core equations of
the LSTM as proposed by (\textbf{hochreiter1997long?}):

\begin{equation}\phantomsection\label{eq-LSTM}{
\begin{aligned}
\vec{I}^t[\tau] &= \sigma(\vec{i}^t[\tau]) \\
\vec{F}^t[\tau] &= \sigma(\vec{f}^t[\tau]) \\
\vec{O}^t[\tau] &= \sigma(\vec{o}^t[\tau]) \\
\vec{C}^t[\tau] &= \vec{F}^t[\tau] \circ \vec{C}^t[\tau-1]  + \vec{I}^t[\tau] \circ \tanh(\vec{c}^t[\tau]) \\
\vec{H}^t[\tau]  &= \vec{O}^t[\tau] \circ \tanh(\vec{C}^t[\tau]) \ ,
\end{aligned}
}\end{equation}

where \(\sigma\) denotes the logistic sigmoid function, \(\tanh\) is the
hyperbolic tangent and the \(\circ\) stands for the Hadamard product
(all applied in an element-wise fashion to the vectors). In the last two
equations the role of the input, forget and output gates as described
above becomes apparent.

The third step in a single layer LSTM (as employed for the work
presented here) is then to provide the output of the current LSTM cell,
i.e.~\(\vec{H}^t[\tau]\) and \(\vec{C}^t[\tau]\), to the subsequent LSTM
cell that processes the next element \(\vec{X}^t[\tau+1]\) of the input
sequence.

The full action of the LSTM network up to the end of the sequence can be
written as a nested function call

\begin{equation}\phantomsection\label{eq-L_nested}{
\left(\vec{H}^t[N_\tau], \vec{C}^t[N_\tau] \right) = L \left( \vec{X}^t[N_\tau], L \left( \vec{X}^t[N_\tau-1], \ldots L \left( \vec{X}^t[1], (\vec{H}^t[0], \vec{C}^t[0]) \right) \ldots \right) \right) \ ,
}\end{equation}

where
\(L\left(\vec{X}^t[\tau], \left(\vec{H}^t[\tau-1], \vec{C}^t[\tau-1] \right) \right)\)
represents Equation~\ref{eq-LSTM-gates} and Equation~\ref{eq-LSTM}. For
the present work, the initial conditions are chosen as
\(\vec{H}^t[0]=\vec{C}^t[0]=0\), which simply means that there is no
memory longer than \(N_\tau\) time steps.

The final output of the ConvLSTM chain, \(\vec{H}^t[N_\tau]\) and
\(\vec{C}^t[N_\tau]\), encode information on the full input sequence
ending at time \(t\). This information has to be decoded via an
appropriate subsequent network, as it is described in the
Section~\ref{sec-FC_layer}.

\subsubsection{Combining the LSTM with spatial
convolution}\label{combining-the-lstm-with-spatial-convolution}

Although the plain LSTM has high performance in handling temporal
sequences of point-like quantities it is not designed to recognize
spatial features in sequences of two-dimensional maps as
atmosphere-ocean interface fields. To address this limitation we employ
a ConvLSTM architecture as described in the following.

In this kind of network the elements of the input sequence are given as
spatially varying fields
\(\vec{X}^t[\tau] = (X^t_k[x,y,\tau]) \in \mathbb{R}^{N_k \times (N_x \times N_y)}\),
where \(x\in[1, N_x]\) and \(y \in [1, N_y]\) run over the horizontal
and vertical dimensions of the map, respectively. In order to enable the
``learning'' of spatial patterns, the free parameters of the network are
replaced by two-dimensional convolution kernels
\(\pmb{\mathcal{M}}^{g} = (\mathcal{M}^{g}_{hk}[\xi, \eta]) \in \mathbb{R}^{(N_h \times N_k)\times (N_\xi \times N_\eta)}\)
and
\(\pmb{\mathcal{N}}^{g} = (\mathcal{N}^{g}_{hh'}[\xi, \eta]) \in \mathbb{R}^{(N_h \times N_h)\times (N_\xi \times N_\eta)}\).
The width and the height of the kernels are given by \(N_\xi\) and
\(N_\eta\), respectively and
\(\xi\in [-(N_\xi-1)/2,(N_\xi-1)/2], \eta\in [-(N_\eta-1)/2,(N_\eta-1)/2]\),
where, without loss of generality, we assume odd numbers for the kernel
sizes.

The mapping from the input quantities to the gates is then given by a
convolution with these kernels \[
\begin{aligned}
g^t_h[x,y,\tau] & =  \mathcal{M}^{g}_{hk} [\xi,\eta]\, X^t_k[x-\xi, y-\eta, \tau]  +  \mathcal{N}^{g}_{hh'}[\xi,\eta] \, H^t_{h'}[x-\xi, y-\eta, \tau-1] + \mathcal{B}^g_{h}\ .
\end{aligned}
\]

again with Einstein's convention imposed.

It becomes immediately apparent that in case of the ConvLSTM, the gate
and state vectors must become vector fields
(\(\in \mathbb{R}^{N_h \times (N_x \times N_y)}\)) as well. We can write
this mapping in the same way as Equation~\ref{eq-LSTM-gates} but with
replacing the normal matrix-vector multiplication by a convolution
(denoted with \(\ast\)), i.e.\\
\[
\begin{aligned}
\vec{g}^t[\tau] = \pmb{\mathcal{M}}^{g} \ast \vec{X}^t[\tau] + \pmb{\mathcal{N}}^{g} \ast  \vec{H}^t[ \tau-1]+ \vec{\mathcal{B}}^g \,
\end{aligned}
\]

The subsequent processing of the \(\vec{g}^t[\tau]\) remains
symbolically the same as presented in Equation~\ref{eq-LSTM} but with
all appearing quantities now meaning vector fields instead of simple
vectors.

In summary, the ConvLSTM is designed for processing tasks that demand a
combined understanding of spatial patterns and temporal sequences in
data. It merges the image-processing capabilities of Convolutional
Neural Networks (CNNs) with the time-series modeling of Long Short-Term
Memory (LSTM) networks.

\subsection{Fully connected layer}\label{sec-FC_layer}

As stated in Section~\ref{sec-LSTM}, the final output
\(\vec{H}^t[N_\tau]\) and \(\vec{C}^t[N_\tau]\) of the ConvLSTM encode
information on the full input sequence. In order to contract this
information to obtain the runoff vector \(\vec{R}^t\) representing the
\(N_r\) rivers, we propose to subject the final short-term memory
(i.e.~the hidden state \(\vec{H}^t[N_\tau]\)) to an additional FC
network.

In particular, the dimensionality of the vector field
\(\vec{H}^t[N_\tau]\) is sequentially reduced by three nested FC layers,
each connected to the other by the Rectified Linear Unit (ReLU), see
Figure~\ref{fig-ConvLSTM_FC}. Integrating out artificial degrees of
freedom in a step-wise fashion has turned out to be beneficial
{[}\texttt{missing}{]}.

Spelled out in mathematics, the runoff of the \(r\)-th river is then
obtained via (using Einstein's convention)

\[
\begin{aligned}
R_r^t = \mathcal{W}^{3}_{rb}\mathrm{ReLU}\left(\mathcal{W}^{2}_{ba}\mathrm{ReLU}\left(\mathcal{W}^{1}_{ah} [x,y] H^t_h[x,y, N_\tau] + \mathcal{B}^1_a\right) + \mathcal{B}^2_b \right) + \mathcal{B}^3_r \ ,
\end{aligned} 
\]

where \(a=1,\ldots N_a\), \(b=1,\ldots N_b\) and the hyper parameters
\(N_a\) and \(N_b\) are chosen such that
\(N_h\cdot N_x\cdot N_y > N_a > N_b > N_r\) in order to achieve the
aforementioned step-by-step reduction of dimensionality. The weights
\(\mathcal{W}\) and biases \(\mathcal{B}\) stand for parameters that are
optimized during the training of the network.

In matrix-vector notation this can be compressed to
\begin{equation}\phantomsection\label{eq-FC_layer}{
\begin{aligned}
\vec{R}^t = \pmb{\mathcal{W}}^{3}\mathrm{ReLU}\left(\pmb{\mathcal{W}}^{2}\mathrm{ReLU}\left(\pmb{\mathcal{W}}^{1} \vec{H}^t[N_\tau] + \vec{\mathcal{B}}^1\right) + \vec{\mathcal{B}}^2 \right) + \vec{\mathcal{B}}^3 \ .
\end{aligned}
}\end{equation}

Combining Equation~\ref{eq-FC_layer} with Equation~\ref{eq-L_nested}
provides finally an explicit formula for the initial assumption of
modelling the runoff for time \(t\) as a functional of a sequence of
atmospheric fields, i.e.

\[
\begin{aligned}
\vec{R}^t & = \vec{M}(\{X^t_k[x,y,\tau]\}) \\
& = \pmb{\mathcal{W}}^{3}\mathrm{ReLU}\left(\pmb{\mathcal{W}}^{2}\mathrm{ReLU}\left(\pmb{\mathcal{W}}^{1} \vec{L}_H \left( \vec{X}^t[N_\tau], L \left( \vec{X}^t[N_\tau-1], \ldots L \left( \vec{X}^t[1], (0, 0) \right) \ldots \right) \right) + \vec{\mathcal{B}}^1\right) + \vec{\mathcal{B}}^2 \right) + \vec{\mathcal{B}}^3 \ ,
\end{aligned}
\] where the \(\vec{L}_H\) means that only the hidden state vector of
the final ConvLSTM call is forwarded to the FC layer.

In Section~\ref{sec-technical_details}, we present the employed model
and data setup as well as the choice for all mentioned hyper parameters
that lead to an adequate model performance after training.

\section{Technical details}\label{sec-technical_details}

\subsection{Runoff data used for
training}\label{runoff-data-used-for-training}

The runoff data covering the period 1979 to 2011 is based on an E-HYPE
hindcast simulation that was forced by a regional downscaling of
ERA-Interim (\textbf{dee2011?}) with RCA3 (\textbf{theross?}) and
implemented into NEMO-Nordic (\textbf{hordoir2019?}) as a mass flux. For
the periods before (1961 to 1978) and after (2012 to 2018) additional
spatial temporal corrections have been applied to the runoff data, and
have therefore been ignored. The quality of the runoff was extensively
evaluated. For more information see (\textbf{gröger2022?}) and
references therein.

\subsection{Atmospheric Forcing}\label{atmospheric-forcing}

The UERRA-HARMONIE regional reanalysis dataset was developed as part of
the FP7 UERRA project (Uncertainties in Ensembles of Regional
Re-Analyses, \href{http://www.uerra.eu/}{http://www.uerra.eu/)},). The
UERRA-HARMONIE system represents a comprehensive, high-resolution
reanalysis covering a wide range of essential climate variables. This
dataset encompasses data on air temperature, pressure, humidity, wind
speed and direction, cloud cover, precipitation, albedo, surface heat
fluxes, and radiation fluxes from January 1961 to July 2019. With a
horizontal resolution of 11 km and analyses conducted at 00 UTC, 06 UTC,
12 UTC, and 18 UTC, it also provides hourly resolution forecast model
data. UERRA-HARMONIE is accessible through the Copernicus Climate Data
Store (CDS, \url{https://cds.climate.copernicus.eu/\#!/home),} initially
produced during the UERRA project and later transitioned to the
Copernicus Climate Change Service (C3S,
\url{https://climate.copernicus.eu/copernicus-regionalreanalysis-europe).}

\subsection{Ocean Model}\label{ocean-model}

To simulate the Baltic Sea, we use a coupled 3-dimensional ocean model
Baltic Sea, called the Modular Ocean Model (MOM). This model uses a
finite-difference method to solve the full set of primitive equations to
calculate the motion of water and the transport of heat and salt. The
K-profile parameterization (KPP) was used as turbulence closure scheme.
The model's western boundary opens into the Skagerrak and connects the
Baltic Sea to the North Sea. The maximum depth was set at 264 meters. A
more detailed description of the setup can be found in
(\textbf{gröger2022?}).

\subsection{Neural network hyper
parameters}\label{neural-network-hyper-parameters}

During training, the following set of hyper parameters has turned out to
yield sufficient quality of the results for the present study:

\begin{longtable}[]{@{}ll@{}}
\caption{Hyperparameters}\label{tbl-letters}\tabularnewline
\toprule\noalign{}
Parameter name & Parameter size \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Parameter name & Parameter size \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Channel size & 4 \\
Num. hidden layer & 9 \\
Num. timesteps & 30 \\
Conv. kernel size & (7,7) \\
Num. ConvLSTM layers & 1 \\
Batch size & 50 \\
Learning Rate & 1e-3 with CosineAnnealing \\
\end{longtable}

As input the model receives \(N_{\tau}\) = 30 days of atmospheric
surface fields temperature, precipitation, specific humidity and wind
speed, with a daily resolution to predict the river runoff \(\vec{R}\)
(Figure~\ref{fig-baltNet}).

\begin{figure}

\centering{

\includegraphics{ConvLSTM.png}

}

\caption{\label{fig-baltNet}Schematic structure of the ConvLSTM
implementation for river runoff forecasting.}

\end{figure}%

The choice of atmospheric fields was based on the assumption that the
runoff should be mainly given by the net precipitation (precipitation
minus evaporation). The evaporation flux is often calculated as a
function of wind speed, the air's humidity, the air's density and the
involved turbulent exchange coefficients (Karsten et al., 2024), where
the latter two are influenced by the air temperature.

\section{Results}\label{results}

\subsection{ConvLSTM model evaluation}\label{sec-evaluation}

The model was trained with daily data for the period 1979 to 2011, as
this period represents the only period of E-HYPE without further bias
correction applied to the runoff to match observations. The data was
divided into randomly chosen splits of 80\(\%\) training data, 10\(\%\)
validation data to evaluate the performance of the model during
training, and 10\(\%\) test data which is finally used to evaluate the
performance of the model after training. The model was trained for 400
epochs and the model weights with the lowest mean squared error during
training have been stored.

The accuracy of the model is displayed in
Figure~\ref{fig-statistical-evaluationNN}. As mention above, for this
evaluation, the model's output is compared to the test data, which,
importantly, the model has not seen yet during the training.
\texttt{Which\ river\ is\ shown\ here?} The left panel a) illustrates
the relative prediction error in relation to the original E-HYPE data,
indicating that, on daily timescales, the model can predict river runoff
with an accuracy of \(\pm\) 5\(\%\). The overall correlation is 0.997
with the resulting error metrics yielding a root mean square error
(RMSE) of 323.99 \(m^3\)/s and mean absolute error (MAE) of 249.51
\(m^3\)/s. While the model's performance is satisfactory, the
discrepancies between the actual values and the predictions could partly
be attributed to the use of a different atmospheric dataset than the one
originally used to drive the E-HYPE model. However, by applying a
rolling mean with a 5-day window, the prediction error is reduced to
less than 1\(\%\), which is acceptable for the purposes of climate
modeling. Lastly, the right panel demonstrates that the distribution of
residuals follows a Gaussian shape, suggesting that our model does not
exhibit bias by systematically over or underestimating river runoff
values.

\begin{figure}

\centering{

\captionsetup{labelsep=none}\includegraphics{../src/figures/paper_error_metrics.png}

}

\caption{\label{fig-statistical-evaluationNN}}

\end{figure}%

\texttt{The\ time\ axis\ should\ be\ improved\ here\ and\ in\ all\ other\ time\ series\ plots.}

In the following the model's performance to reproduce the total river
runoff as well as the inflow of the four largest rivers into the Baltic
Sea is addressed. Figure~\ref{fig-PerformanceNeuralNetworkRunoff} shows
the predicted and the original river runoff in comparison to the test
dataset. The predicted total river runoff for the Baltic Sea is closely
matching the original data, see panel a) therin. Zooming in on the
largest individual rivers (panels b-e), it can be seen that also the
prediction of the individual rivers is close to the original data.

\begin{figure}

\centering{

\captionsetup{labelsep=none}\includegraphics{../src/figures/paper_total_river_runoff.png}

}

\caption{\label{fig-PerformanceNeuralNetworkRunoff}}

\end{figure}%

\subsection{Application of the ConvLSTM in combination with an ocean
model}\label{sec-MOM-application}

Lastly, we evaluate the performance of the ConvLSTM by incorporating the
predicted river runoff as hydrological forcing into the ocean model
MOM5. This provides a robust validation of the runoff model against more
complex real world conditions and is thereby ensuring that the
predictions accurately reflect the impact of the river discharge on the
ocean dynamics. This in turn validates that the temporal and spatial
variability of river runoff is well captured by ConvLSTM and that the
residuals shown in Figure~\ref{fig-statistical-evaluationNN} are indeed
insignificant when it comes to realistic applications.

Figure~\ref{fig-by15} shows the salinity comparison between the original
E-HYPE river runoff and the predicted river runoff at BY15 - a central
station in the Baltic Sea, east of the Gotland island. It can be seen
that the model simulation using the predicted river runoff by the
ConvLSTM is closely mirroring the control simulation, that is forced
with the E-HYPE runoff. The upper panel a) shows the surface salinity,
representing the high-frequency variations in the salinity, which is
heavily affected by river runoff. The lower panel b) shows the bottom
salinity which in turn is low-frequency feature in the Baltic Sea that
is also well reproduced with the ConvLSTM predictions.

\begin{figure}

\centering{

\captionsetup{labelsep=none}\includegraphics{../src/figures/by15_model.png}

}

\caption{\label{fig-by15}}

\end{figure}%

The final evaluation of the ConvLSTM model concentrates on the spatial
accuracy of river runoff predictions as visualized in
Figure~\ref{fig-evaluation_MOM}. Panel a) therein exhibits the
vertically averaged salinity for the period 1981 to 2011 in the
reference simulation. It highlights the strong horizontal gradients and
complex topographic features in the Baltic Sea, as evidenced by salinity
variations in deeper waters, captured by the vertical integration. In
panel b), these reference results are compared to the ConvLSTM
simulation by showing the percentage difference in vertically averaged
salinity. Overall, the differences remain below 1\(\%\), except in the
Gulf of Riga (22-24°E, 56.5-58.5°N for orientation), where the runoff is
dominated by the Daugava river. Here the difference is approximately
1\(\%\).

\begin{figure}

\centering{

\captionsetup{labelsep=none}\includegraphics{../src/figures/evaluation_MOM.png}

}

\caption{\label{fig-evaluation_MOM}}

\end{figure}%

\section{Discussion}\label{sec-discussion}

With the increasing demand of decision makers for regional climate
projections, that allow to quantify regional climate change impacts, the
availability of precise hydrological forecasting becomes invaluable. The
quality of a projection for a coastal sea such as the Baltic Sea is too
large parts based on the quality of the hydrological conditions. In this
work we analyze the implementation of a ConvLSTM networks for predicting
river runoff, highlighting its potential to enhance river runoff
forecasting across different coastal seas. Given the unique hydrological
characteristics of the Baltic Sea, largely influenced by its limited
connection to the open ocean and significant freshwater input from
surrounding rivers, the region presents a critical case for the
application of sophisticated forecasting models.

The transition from traditional hydrological models to machine learning
approaches, such as the ConvLSTM model, offers significant advantages
with respect to the following aspects. The ConvLSTM can be seamlessly
integrated into regional climate models, allowing for the real-time
computation of river runoff while performing climate projections. While
the initial training of the model requires substantial computational
resources, it remains less intensive than running comprehensive
hydrological models. Furthermore, once trained, the ConvLSTM model is
computationally efficient during inference, ensuring enhanced
forecasting capabilities without significantly increasing computational
demands.

However, while the ConvLSTM represents an advancement for the climate
community, given that running and tuning traditional hydrological models
demands extensive expertise, models like E-HYPE maintain an essential
role. They provide a comprehensive dataset that enables to train our
ConvLSTM model effectively. This robust training helps the machine
learning models to achieve highly accurate predictive weights. Thus,
rather than rendering traditional methods obsolete, the integration of
machine learning models builds upon and enhances the foundational data
provided by them.

\section{Conclusions}\label{sec-conclusions}

In summary, we showed that the ConvLSTM model demonstrated robust
performance in forecasting river runoff across 97 rivers entering the
Baltic Sea. Trained on data from 1979 to 2011, the model achieved an
impressive daily prediction accuracy of \(\pm 5 \%\). Importantly, this
accuracy is achieved without any expert's domain knowledge on
hydrological modelling. This capability to generate accurate and
detailed simulations enables to examine the potential impacts of
different climate scenarios. Such precision in forecasting and scenario
testing is crucial for crafting effective water resource management
strategies and adapting to the changing climate.

The robust performance of the ConvLSTM model in simulating river runoff
and its possible effective integration into coupled regional climate
models, as for example in the IOW ESM (Karsten et al., 2024), pave the
way for a multitude of new storyline simulations. Hence, we can now
explore various ``what-if'' scenarios more reliably, under the
assumption that the model weights attained during training are robust.
This represents a significant step forward in our ability to understand
and predict the complex dynamics of river systems and their impact on
regional climate systems.

\section{Acknowledgments}\label{acknowledgments}

The research presented in this study is part of the Baltic Earth program
(Earth System Science for the Baltic Sea region, see
\href{https://www.baltic.earth/}{https://www.baltic.earth}.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\vspace{1em}

\bibitem[\citeproctext]{ref-fangExaminingApplicabilityDifferent2019}
Fang, W., Huang, S., Ren, K., Huang, Q., Huang, G., Cheng, G., \& Li, K.
(2019). Examining the applicability of different sampling techniques in
the development of decomposition-based streamflow forecasting models.
\emph{Journal of Hydrology}, \emph{568}, 534--550.
\url{https://doi.org/10.1016/j.jhydrol.2018.11.020}

\bibitem[\citeproctext]{ref-hagemannHighResolutionDischarge2020}
Hagemann, S., Stacke, T., \& Ho-Hagemann, H. T. M. (2020). High
{Resolution Discharge Simulations Over Europe} and the {Baltic Sea
Catchment}. \emph{Frontiers in Earth Science}, \emph{8}.
\url{https://doi.org/10.3389/feart.2020.00012}

\bibitem[\citeproctext]{ref-karsten2024}
Karsten, S., Radtke, H., Gröger, M., Ho-Hagemann, H. T. M., Mashayekh,
H., Neumann, T., \& Meier, H. E. M. (2024). Flux coupling approach on an
exchange grid for the IOW earth system model (version 1.04.00) of the
baltic sea region. \emph{Geoscientific Model Development}, \emph{17}(4),
1689--1708. \url{https://doi.org/10.5194/gmd-17-1689-2024}

\bibitem[\citeproctext]{ref-neumann2021}
Neumann, T., Koponen, S., Attila, J., Brockmann, C., Kallio, K.,
Kervinen, M., et al. (2021). Optical model for the baltic sea with an
explicit CDOM state variable: A case study with model ERGOM (version
1.2). \emph{Geoscientific Model Development}, \emph{14}(8), 5049--5062.
\url{https://doi.org/10.5194/gmd-14-5049-2021}

\bibitem[\citeproctext]{ref-shiConvolutionalLSTMNetwork2015}
SHI, X., Chen, Z., Wang, H., Yeung, D.-Y., Wong, W., \& WOO, W. (2015).
Convolutional {LSTM Network}: {A Machine Learning Approach} for
{Precipitation Nowcasting}. In \emph{Advances in {Neural Information
Processing Systems}} (Vol. 28). Curran Associates, Inc. Retrieved from
\url{https://proceedings.neurips.cc/paper/2015/hash/07563a3fe3bbe7e3ba84431ad9d055af-Abstract.html}

\bibitem[\citeproctext]{ref-tanAdaptiveMiddleLongterm2018}
Tan, Q.-F., Lei, X.-H., Wang, X., Wang, H., Wen, X., Ji, Y., \& Kang,
A.-Q. (2018). An adaptive middle and long-term runoff forecast model
using {EEMD-ANN} hybrid approach. \emph{Journal of Hydrology},
\emph{567}, 767--780.
\url{https://doi.org/10.1016/j.jhydrol.2018.01.015}

\end{CSLReferences}

\end{document}